Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic: adapted to CHOP input format

### 2022-01-21 psy_peds_3_CHOP.Rmd
Integrate additional analysis to the CHOP pipeline. Analysis by disorder group

listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "lubridate")
paket(listOfPackages)
```

@@

```{r message=FALSE, warning=FALSE}
outputDir <- "../output/CHOP/"
allCounts <- read.csv( file = paste0( outputDir, "All-Counts.csv"))
psyCounts <- read.csv( file = paste0( outputDir, "Psyc-Counts.csv"))

@@

```{r}
# Time series analysis of percentage of psy codes per encounter, linear (without seasonality)
patient_count_psy_period_psy <- ratios_patients_with_without_psy %>%
  filter(time_p > as.Date(cut( start_date_plots, breaks = time_period)) &
           time_p < as.Date(cut( end_date_plots, breaks = time_period))) %>%
  mutate(time = seq_along(time_p),
         month_year = format(time_p, "%m")
  )

@@

>>>>

listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "lubridate", "magrittr")
paket(listOfPackages)
```

@@

```{r message=FALSE, warning=FALSE}
allCounts <- read.csv( "../output/CHOP/All-Counts.csv")
psyCounts <- read.csv( "../output/CHOP/Psyc-Counts.csv")

@@

```{r}
# Time series analysis of percentage of psy codes per encounter, linear (without seasonality)
patient_count_psy_period_psy <- left_join(allCounts,
                                          psyCounts) %>%
  mutate(count_no_psy_patients = count_all - count_psy_patients,
         time = seq_along(time_p),
         period = ifelse(time_p < as.Date("2020-04-01"),
                         "before_pandemic",
                         "during_pandemic"),
         ratio = count_psy_patients / count_no_psy_patients,
         percentage = count_psy_patients * 100 / (count_psy_patients + count_no_psy_patients),
         month_year = format.Date(time_p, "%m"),
         siteid = "CHOP") %>%
  select(time_p, count_no_psy_patients, period, count_psy_patients,
         ratio, percentage, time, month_year, siteid)

@@

## Psychiatric conditions aggregating by subcategories
```{r}
# Generate CHOP tables aggregated by sex and by mental disorder category
psy_female_counts <- read_csv("../output/CHOP/Psy-Female-11-17.csv")
psy_male_counts <- read_csv("../output/CHOP/Psy-Male-11-17.csv")

psy_tot <- psy_female_counts %>% 
  select(c(-Month, -Year)) %>%
  add( psy_male_counts %>% select(c(-Month, -Year))) %>%
  mutate(time_p = as.Date(paste(psy_female_counts$Year, psy_female_counts$Month, "01", sep = "-")))
```

```{r}
count_disorder_group <- psy_tot %>%
  pivot_longer(!time_p, names_to = "disorder_group", values_to = "count_dg") %>%
  inner_join( patient_count_psy_period_psy %>%
                mutate( total = count_no_psy_patients + count_psy_patients) %>%
                select( time_p, count_no_psy_patients, count_psy_patients, total, period ))

dg <- count_disorder_group %>%  
  ggplot(aes(x = time_p, y= count_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = time_period,
       y = "Patient Count",
       title = paste0("Per ", time_period, " patient counts with mental health related ICD codes"), 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])
dg
```

#### Subset: the most prevalent disorder group
```{r}
dg_subset <- count_disorder_group %>%
  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
  ggplot(aes(x = time_p, y= count_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = time_period,
       y = "Patient Count",
       title = paste0("Per ", time_period, " patient counts with mental health related ICD codes"), 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 12),
        axis.text.x = element_text(size = 12, angle = 90), 
        axis.text.y = element_text(size = 11), 
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 11),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])
dg_subset
```

### New admissions: percentage

```{r}
# estimate the percentage using as denominator the total number of patients per month with any psychiatric condition
perc_disorder_group <- count_disorder_group %>%
  mutate( percentage_dg = 100*count_dg / count_psy_patients )

dgPerc <- perc_disorder_group %>%  
  ggplot(aes(x = time_p, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = time_period,
       y = "Patient (%)",
       title = paste0("Percentage of patients per ", time_period,"with mental health related ICD codes"), 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])
dgPerc
```

```{r}
dgPerc_subset <- perc_disorder_group %>%  
  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
  ggplot(aes(x = time_p, y= percentage_dg, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = time_period,
       y = "Patient (%)",
       title = paste0("Percentage of patients per ", time_period,"with mental health related ICD codes"), 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 12),
        axis.text.x = element_text(size = 12, angle = 90), 
        axis.text.y = element_text(size = 12), 
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])
dgPerc_subset
```

# Save the output in a RData file. 
```{r}
site = "CHOP"
patient_count_psy_period_psy$siteid = site 
patient_count_psy_period_psy_clear$siteid = site 
bootstrapped_coefficients_df$siteid = site 
bootstrapped_fitted_df$siteid = site 
ratios_patients_with_without_psy$siteid = site 
quantile_bootstrapped_fitted$siteid = site
perc_disorder_group$siteid = site

save( ratios_patients_with_without_psy,
      patient_count_psy_period_psy,
      patient_count_psy_period_psy_clear,
      bootstrapped_coefficients_df,
      bootstrapped_fitted_df,
      quantile_bootstrapped_fitted,
      perc_disorder_group,
      file = paste0("../output/CHOP/peds_psy_aggResults_", site, ".RData"))
